Revisiting the Classics: Online RL in the Programmable Dataplane

### 2021-06-03 ngni-talk/talk.tex
NGNI talk draft.

\usepackage{bm}

@@

@@

\title{Towards In-Switch Reinforcement Learning}
\author{\vspace{-1em}\textbf{Kyle A. Simpson}\\
	\faEnvelopeO{} \href{mailto:k.simpson.1@research.gla.ac.uk}{\nolinkurl{k.simpson.1@research.gla.ac.uk}}\\
	\vspace{1em}\small{\faGithub{} \href{https://github.com/felixmcfelix}{FelixMcFelix} \hspace{0.5em} \faGlobe{} \url{https://mcfelix.me}}}
\institute{University of Glasgow}
\date{\nth{2} December, 2020}

\begin{document}
\begin{frame}
\maketitle
\end{frame}

@@

\begin{frame}{Case study: DDoS Prevention}
	\begin{columns}
		\begin{column}{0.4\linewidth}
			\begin{figure}
				\centering
				\resizebox{\linewidth}{!}{
					\begin{tikzpicture}[
					texts/.style = {text=black},
					labeltexts/.style = {text=gray},
					treeline/.style = {draw=uofgburgundy},
					treenode/.style = {texts, circle, centered, fill=white, treeline},
					load/.style = {fill=uofgcobalt},
					loadline/.style = {draw=uofgcobalt, line width=0.75mm},
					loadhide/.style = {fill=uofgcobalt!40!white},
					external/.style = {fill=uofgrust},
					externalhide/.style = {fill=uofgrust!40!white},
					hideline/.style = {draw=uofgsandstone!40!white},
					hidenode/.style = {treenode, hideline},
					grow'=right
					]
					\node [treenode, loadhide, label={[texts]above:Agent 1}] (mainagent) {};
					\node [treenode, loadhide, right = of mainagent] (inner0) {};
					\node [treenode, loadhide, above right = 0.2cm and 1cm of inner0] (inner1) {};
					\node [hidenode, below right = 0.2cm and 1cm of inner0] (inner2) {};
					\node [treenode, loadhide, below right = 0.2cm and 1cm of inner1] (inner3) {};
					\node [treenode, loadhide, right = of inner3] (inner4) {};
					\node [treenode, loadhide, above right = 0.2cm and 1cm of inner4, label={[texts]above:$s_0$}] (s0) {};
					\node [hidenode, below right = 0.2cm and 1cm of inner4, label={[labeltexts]above:$s_1$}] (s1) {};

					\node [hidenode, above left = 0.2cm and 0.5cm of inner1, label={[labeltexts]above:Agent 2}] (agent2) {};
					\node [hidenode, below left = 0.2cm and 0.5cm of inner2, label={[labeltexts]below:Agent 3}] (agent3) {};

					\draw[hideline, -] ($ (mainagent) + (-1,0) $) -- (mainagent);
					\draw[hideline, -] ($ (agent2) + (-1,0) $) -- (agent2);
					\draw[hideline, -] ($ (agent3) + (-1,0) $) -- (agent3);
					\draw[hideline, -] (inner1) -- (agent2);
					\draw[hideline, -] (inner2) -- (agent3);

					\draw[loadline, -] (mainagent) -- (inner0);
					\draw[loadline, -] (inner0) -- (inner1);
					\draw[hideline, -] (inner0) -- (inner2);
					\draw[-] (inner1) -- (inner3);
					\draw[hideline, -] (inner2) -- (inner3);
					\draw[loadline, -] (inner3) -- (inner4);
					\draw[loadline, -] (inner4) -- node [texts, above] {$U_{s_0}$} (s0) ;
					\draw[hideline, -] (inner4) -- node [labeltexts, below] {$U_{s_1}$} (s1);

					\end{tikzpicture}
				}
			\caption{`Global' state for any flow (\textbf{bold} means used by policy).}
			\end{figure}
		\end{column}
		\begin{column}{0.6\linewidth}
			\begin{itemize}
				\item Prior work\footnotemark.
				\item Monitor flow statistics at ingress/egress points.
				\item Use load measurements from flow paths as global state.
				\item OUTPUT: throttle individual flows.
				\item Classical RL.
				\begin{itemize}
					\item Learning time on order of minutes.
				\end{itemize}
			\end{itemize}
		\end{column}
	\end{columns}
\footcitetext{DBLP:journals/tnsm/SimpsonRP20}
\end{frame}

\begin{frame}{Case Study: DDoS Prevention (Architecture)}
	\begin{columns}
		\begin{column}{0.3\linewidth}
			\begin{itemize}
				\item Reward measurements come from network.
				\item Input state mixes local flow data with global load data.
				\item Flow measurements combined, decisions scheduled to prevent overload.
			\end{itemize}
		\end{column}
		\begin{column}{0.7\linewidth}
			\begin{figure}
				\centering
				\resizebox{0.7\linewidth}{!}{\begin{tikzpicture}[scale=1, every node/.style={scale=0.7}]
					\node(remote){Remote};

					%%%

					\node[below=0.05 of remote](swpos){};
					\node[fill=white!80!uofgcobalt, draw=black, minimum height=1.5cm, minimum width=2cm, below right= 0.1 of swpos.north west](sw1){};
					\node[fill=white!80!uofgcobalt, draw=black, minimum height=1.5cm, minimum width=2cm, below right= 0.1 of sw1.north west](sw2){};
					\node[fill=white!80!uofgcobalt, draw=black, minimum height=1.5cm, minimum width=2cm, below right= 0.1 of sw2.north west](switch){};
					\node[below right, inner sep=2pt] at (switch.north west) {\small Switch};
					\node[fill=white!90!uofgcobalt, draw, rectangle, rounded corners=0.05cm, above=0.1] (oswlc) at (switch.south) {\begin{varwidth}{1.5 cm}\scriptsize \centering Load\\Collector\end{varwidth}};

					%

					\node[right=2 of swpos](epos){};
					\node[fill=white!80!uofgthistle, draw=black, minimum height=1.5cm, minimum width=3.5cm, below right= 0.1 of epos.north west](e1){};
					\node[fill=white!80!uofgthistle, draw=black, minimum height=1.5cm, minimum width=3.5cm, below right= 0.1 of e1.north west](e2){};
					\node[fill=white!80!uofgthistle, draw=black, minimum height=1.5cm, minimum width=3.5cm, below right= 0.1 of e2.north west](egress){};
					\node[below right, inner sep=2pt] at (egress.north west) {\small Egress Switch};
					\node[fill=white!90!uofgthistle, draw, rectangle, rounded corners=0.05cm, above=0.1] (eglc) at ($(egress.south) + (-0.65,0)$) {\begin{varwidth}{1.5 cm}\scriptsize \centering Load\\Collector\end{varwidth}};
					\node[fill=white!90!uofgthistle, draw, rectangle, rounded corners=0.05cm, right=0.1] (egest) at (eglc.east) {\begin{varwidth}{1.5 cm}\scriptsize \centering Estimator\\$g(\cdot)$\end{varwidth}};

					%

					\node[right=3.5 of epos](apos){};
					\node[fill=white!60!uofgpumpkin, draw=black, minimum height=1.5cm, minimum width=2cm, below right= 0.1 of apos.north west](a1){};
					\node[fill=white!60!uofgpumpkin, draw=black, minimum height=1.5cm, minimum width=2cm, below right= 0.1 of a1.north west](a2){};
					\node[fill=white!60!uofgpumpkin, draw=black, minimum height=1.5cm, minimum width=2cm, below right= 0.1 of a2.north west](otheragent){};
					\node[below right, inner sep=2pt] at (otheragent.north west) {\small Agent vNF};
					\node[fill=white!90!uofgpumpkin, draw, rectangle, rounded corners=0.05cm, above=0.3] (oasa) at (otheragent.south) {\begin{varwidth}{1.5 cm}\small \centering Stats API\end{varwidth}};

					%

					\node[below=2.3 of remote.west](linestart){};
					\path let \p1 = (linestart) in node (lineend) at (9,\y1){};
					\draw [dashed] (linestart) -- (lineend);

					%%%

					\node[below=2.4 of remote](local){Local};

					%%%

					\node[below=0.05 of local](aswpos){};
					\node[fill=white!80!uofgthistle, draw=black, minimum height=3.5cm, minimum width=2.2cm, below right= 0.1 of aswpos.north west](aswitch){};
					\node[below right, inner sep=2pt] at (aswitch.north west) {\small Agent Switch};
					\node[fill=white!90!uofgthistle, draw, rectangle, rounded corners=0.05cm, above=0.1] (aswoft) at (aswitch.south) {\begin{varwidth}{1.5 cm}\scriptsize \centering OpenFlow\\Tables\end{varwidth}};
					\node[fill=white!90!uofgthistle, draw, rectangle, rounded corners=0.05cm, above=0.1] (aswsc) at (aswoft.north) {\begin{varwidth}{1.5 cm}\scriptsize \centering Stats\\Collector\end{varwidth}};
					\node[fill=white!90!uofgthistle, draw, rectangle, rounded corners=0.05cm, above=0.1] (aswlc) at (aswsc.north) {\begin{varwidth}{1.5 cm}\scriptsize \centering Load\\Collector\end{varwidth}};

					%

					\node[right=3 of aswpos](avfpos){};
					\node[fill=white!60!uofgpumpkin, draw=black, minimum height=3.5cm, minimum width=4.5cm, below right= 0.1 of avfpos.north west](avf){};
					\node[below right, inner sep=2pt] at (avf.north west) {\small Agent vNF};
					\node[fill=white!90!uofgpumpkin, draw, rectangle, rounded corners=0.05cm, below=0.15] (avfsa) at ($(avf.north) + (0.2,0)$) {\begin{varwidth}{1.5 cm}\small \centering Stats API\end{varwidth}};
					\node[fill=white!90!uofgpumpkin, draw, rectangle, rounded corners=0.05cm, below=0.9] (avfdb) at (avfsa.south west) {\begin{varwidth}{1.5 cm}\scriptsize \centering Flowstate\\Database\end{varwidth}};
					\node[fill=white!90!uofgpumpkin, draw, rectangle, rounded corners=0.05cm, right=0.1] (avfsched) at (avfdb.east) {\begin{varwidth}{1.5 cm}\scriptsize \centering TRS\\Scheduler\end{varwidth}};
					\node[fill=white!90!uofgpumpkin, draw, rectangle, rounded corners=0.05cm, below=2.3] (avfcore) at (avfsa.south) {\begin{varwidth}{1.5 cm}\small \centering Core\end{varwidth}};
					\node[fill=white!90!uofgpumpkin, draw, rectangle, rounded corners=0.05cm, right=0.8] (avfrl) at (avfcore.east) {\begin{varwidth}{1.5 cm}\small \centering RL\end{varwidth}};

					%%%

					\tikzset{>=stealth}

					\draw[thick, ->] (aswlc) -- (avfsa.west) node[midway,above] {\tiny Current load};
					\draw[thick, <->] (aswsc) -- (avfsa.west) node[midway,sloped, above] {\tiny Flow stats};

					\draw[thick, ->] (oswlc) -- (avfsa) node[midway,above, sloped] {\tiny Current load};

					\draw[thick, ->] (eglc) -- (avfsa) node[midway,above, sloped] {\tiny Current load};
					\draw[thick, ->] (egest) -- (avfsa) node[midway,above, sloped] {\tiny Estimation data};

					\draw[thick, <->] (oasa) -- (avfsa) node[midway,above, sloped] {\tiny Experience};

					\draw[thick, ->] (avfsa) -- (avfdb) node[midway,above, sloped] {\tiny State};
					\draw[thick, ->] (avfsa) -- (avfsched) node[midway,above, sloped] {\tiny Live flows};

					\draw[thick, ->] (avfcore) -- (aswoft) node[midway,above, sloped] {\tiny Packet drop rules};
					\draw[thick, <->] (avfcore) -- (avfdb) node[midway,below, sloped] {\tiny State};
					\draw[thick, <->] (avfcore) -- (avfsched) node[midway,below, sloped] {\tiny Work};
					\draw[thick, <-] ($(avfcore.east) + (0,0.1)$) -- ($(avfrl.west) + (0,0.1)$) node[midway,above, sloped] {\tiny Actions};
					\draw[thick, ->] (avfcore) -- (avfrl) node[midway,below, sloped] {\tiny State};

					\draw[thick, ->] (avfsa.east) to [out=0, in=45] (avfrl.north);
					\node[right=0.3] at (avfsa.east) {\tiny Experience};
					\end{tikzpicture}}
				\caption{
					(Non-PDP) system architecture for RL-driven DDoS defence.
					\label{fig:sys-arch}
				}
				\vspace{-1em}
			\end{figure}
		\end{column}
	\end{columns}
\end{frame}

\begin{frame}{Goals... and Reality}
	\centering
	\begin{tabular}{c c c}
		\toprule
		Goal & Solution & Drawback \\ \midrule
		$\bullet$ On-NIC action compute & \onslide+<2->{\alert<2>{Quantisation}} & \onslide+<2->{\alert<2>{Platform Model}} \\
		$\bullet$ On-NIC learning & \onslide+<3->{\alert<3>{Classical RL}} & \onslide+<3->{\alert<3>{Info capacity}} \\
		$\bullet$ Direct action installation & \onslide+<4->{\alert<4>{Custom actions}} & \onslide+<4->{\alert<4>{No intrinsic support}} \\
		$\bullet$ (Partly) External State, Rewards, & \onslide+<5->{\alert<5>{Independent Module}} & \onslide+<5->{\alert<5>{---}} \\
		and Task Independence & & \\
		$\bullet$ Runtime reconfiguration & \onslide+<6->{\alert<6>{P4 Parser + RTE}} & \onslide+<6->{\alert<6>{---}} \\ \bottomrule
	\end{tabular}
\end{frame}

\begin{frame}{Device classes}
	\begin{columns}
		\begin{column}{0.5\linewidth}
			\alert{SmartNICs (e.g., Netronome)}

			\begin{itemize}[<+->]
				\item Low port density.
				\item Easy to have asynchrony:
				\item SOC-based $\rightarrow$ extra cores off the datapath.
				\item NetFPGA $\rightarrow$ can define extra functional units and interconnect.
			\end{itemize}
		\end{column}
		\begin{column}{0.5\linewidth}
			\alert{Programmable Switches (e.g., Tofino)}

			\begin{itemize}[<+->]
				\item High port density.
				\item No additional compute units. Very close to P4 PSA.
				\item But still more powerful--e.g., Tofino supports MUL operations.
				\item Might be doable... deadline-aware.
			\end{itemize}
		\end{column}
	\end{columns}

	\pause
	\vspace{1em}
	We'll mainly focus on SmartNICs.
\end{frame}

\begin{frame}{Architecting On-NIC RL}
	\begin{columns}
		\begin{column}{0.6\linewidth}
			\begin{figure}
				\resizebox{1\linewidth}{!}{\begin{tikzpicture}
						\node[fill=white!80!uofgcobalt, draw=black, minimum height=5.5cm, minimum width=2.9cm, below right= 0.1 of sw2.north west](p4-master){};
						\node[below right, inner sep=2pt] at (p4-master.north west) {\small P4 Datapath};
						\node[fill=white!90!uofgcobalt, draw, rectangle, rounded corners=0.05cm, below=0.8] (p4-rules) at (p4-master.north) {\begin{varwidth}{2.5 cm}\scriptsize \centering Flow Control Rules\end{varwidth}};
						\node[fill=white!90!uofgcobalt, draw, rectangle, rounded corners=0.05cm, below=0.1] (n1) at (p4-rules.south) {\begin{varwidth}{2.5 cm}\scriptsize \centering Extern-based Action\end{varwidth}};
						\node[fill=white!90!uofgcobalt, draw, rectangle, rounded corners=0.05cm, below=0.1] (n2) at (n1.south) {\begin{varwidth}{2.5 cm}\scriptsize \centering ...\end{varwidth}};
						\node[fill=white!90!uofgcobalt, draw, rectangle, rounded corners=0.05cm, below=0.1] (oswlc) at (n2.south) {\begin{varwidth}{1.5 cm}\scriptsize \centering Control Packet Match\end{varwidth}};
						\node[fill=white!90!uofgcobalt, draw, rectangle, rounded corners=0.05cm, below=0.1] (n3) at (oswlc.south) {\begin{varwidth}{2.5 cm}\scriptsize \centering ...\end{varwidth}};
						\node[fill=white!90!uofgcobalt, draw, rectangle, rounded corners=0.05cm, below=0.1] (n4) at (n3.south) {\begin{varwidth}{2.5 cm}\scriptsize \centering Network Telemetry\end{varwidth}};

						\node (pkt-text) at ($(p4-master.north) + (0.0,0.6)$) {\begin{varwidth}{2.5 cm}\scriptsize \centering Packet In\end{varwidth}};

						\node (pkt-text-out) at ($(p4-master.south) + (0.0,-0.6)$) {\begin{varwidth}{2.5 cm}\scriptsize \centering Packet Out\end{varwidth}};

						\node[fill=white!80!uofgsunshine, draw=black, minimum height=5cm, minimum width=5cm](rli) at ($(p4-master.east) + (5, -1)$){};
						\node[below right, inner sep=2pt] at (rli.north west) (rli-text) {\small RL Island};

						\node[draw=black, dotted, minimum height=1cm, minimum width=2cm](wq-out) at ($(rli-text.south) + (0.5,-0.5)$){};
						\node[below right, inner sep=2pt] at (wq-out.north west) {\small EMEM};
						\node[fill=white!90!uofgsunshine, draw, rectangle, rounded corners=0.05cm, above=0.1] (wq-in) at (wq-out.south) {\begin{varwidth}{2.5 cm}\scriptsize \centering Input Queue\end{varwidth}};

						\node[fill=white!90!uofgsunshine, draw, rectangle, rounded corners=0.05cm, above=0.1] (loop) at ($(wq-out.east) + (1.5,-0.5)$) {\begin{varwidth}{2.5 cm}\scriptsize \centering Control ME\end{varwidth}};

						\node[draw=black, dotted, minimum height=1cm, minimum width=2cm](wq-act) at ($(loop.north) + (0.0,1.5)$){};
						\node[below right, inner sep=2pt] at (wq-act.north west) {\small EMEM};
						\node[fill=white!90!uofgsunshine, draw, rectangle, rounded corners=0.05cm, above=0.1] (wq-actin) at (wq-act.south) {\begin{varwidth}{2.5 cm}\scriptsize \centering Action Queue\end{varwidth}};

						\node[draw=black, dotted, minimum height=0.5cm, minimum width=2cm](policy-cls-out) at ($(loop.south) + (0,-1.5)$){};
						\node[below right, inner sep=2pt] at (policy-cls-out.north west) {\small CLS};
						\node[above right, inner sep=2pt] at (policy-cls-out.north west) {\small Policy};
						\node[draw=black, dotted, minimum height=0.5cm, minimum width=2cm](policy-ctm-out) at ($(policy-cls-out.south) + (0,-0.25)$){};
						\node[below right, inner sep=2pt] at (policy-ctm-out.north west) {\small CTM};
						\node[draw=black, dotted, minimum height=0.5cm, minimum width=2cm](policy-emem-out) at ($(policy-ctm-out.south) + (0,-0.25)$){};
						\node[below right, inner sep=2pt] at (policy-emem-out.north west) {\small EMEM};

						\node[fill=white!90!uofgsunshine, draw, rectangle, above=0.1, below left= 0.05] (t1) at (policy-cls-out.north east) {\begin{varwidth}{2.5 cm}\scriptsize \centering T1\end{varwidth}};
						\node[fill=white!90!uofgsunshine, draw, rectangle, above=0.1, below left= 0.05] (t2) at (policy-ctm-out.north east) {\begin{varwidth}{2.5 cm}\scriptsize \centering T2\end{varwidth}};
						\node[fill=white!90!uofgsunshine, draw, rectangle, above=0.1, below left= 0.05] (t3) at (policy-emem-out.north east) {\begin{varwidth}{2.5 cm}\scriptsize \centering T3\end{varwidth}};

						\node[fill=white!90!uofgsunshine, draw, rectangle, rounded corners=0.05cm, above=0.1] (po-me-1) at ($(wq-out.south) + (0,-1.0)$) {\begin{varwidth}{2.5 cm}\scriptsize \centering Policy ME\end{varwidth}};

						\node[fill=white!90!uofgsunshine, draw, rectangle, rounded corners=0.05cm, above=0.1] (po-me-2) at ($(po-me-1.south) + (0,-1.0)$) {\begin{varwidth}{2.5 cm}\scriptsize \centering Policy ME\end{varwidth}};

						\node[fill=white!90!uofgsunshine, draw, rectangle, rounded corners=0.05cm, above=0.1] (po-me-3) at ($(po-me-2.south) + (0,-1.0)$) {\begin{varwidth}{2.5 cm}\scriptsize \centering Policy ME\end{varwidth}};

						\draw[->] (oswlc.east)  to node[midway, above, sloped] {\tiny Control, Requests} (wq-in.west);
						\draw[->] (n4.east)  to node[midway, above, sloped] {\tiny State} (wq-in.west);
						\draw[<->, dotted] (wq-in.east)  to (loop.west);
						\draw[->, dotted] (loop.south) to (policy-cls-out);

						\draw[->] (loop.north)  to node[midway, above, sloped] {\tiny Actions} (wq-actin.south);

						\draw[->] (wq-actin.west)  to node[midway, above, sloped] {\tiny Actions} (n1.east);

						\draw[<->, dotted] (po-me-1.east) to (policy-cls-out.west);
						\draw[<->, dotted] (po-me-2.east) to (policy-ctm-out.west);
						\draw[<->, dotted] (po-me-3.east) to (policy-emem-out.west);

						\draw[->] (pkt-text)  to  (p4-master.north);
						\draw[->] (p4-master.south)  to  (pkt-text-out);

						\draw[->] (loop) to node[midway, above, sloped] {\tiny NN} (po-me-1);
						\draw[->] (po-me-1) to node[midway, above, sloped] {\tiny NN} (po-me-2);
						\draw[->] (po-me-2) to node[midway, above, sloped] {\tiny NN} (po-me-3);
						\draw[->] (po-me-3) to [out=-45,in=0,looseness=2] node[midway, above, sloped] {\tiny NN} (loop);
				\end{tikzpicture}}
				\caption{Architecture for generalised RL agent on Netronome hardware.}
			\end{figure}
		\end{column}
		\begin{column}{0.4\linewidth}
			\begin{itemize}
				\item Classical RL built on tile-coding---\alert{online}.
				\item Async wrt. datapath.
				\item Dynamic selection of last reward, trace info.
				\item \alert{Runtime configurable} (policy, size, application) from data/control-plane.
				\item \alert{Task independent.}
			\end{itemize}
		\end{column}
	\end{columns}
\end{frame}

\begin{frame}{How could this (hypothetically) fit the case study?}
	\begin{itemize}
		\item P4 table-action to gather state on collected flows.
		\begin{itemize}
			\item Flow telemetry in P4 well-documented.
			\item Pass in/schedule state vectors every $\delta{}t$.
		\end{itemize}
		\item Directly pass reward measurements to RL core.
		\item Base policy installed by controller, updated live.
		\item Matched packets poll actions from RL core.
		\begin{itemize}
			\item Map RL actions to state machine, maintain throttling hash table.
			\item Every $\delta{}t'$, batch actions to controller.
		\end{itemize}
	\end{itemize}
\end{frame}

@@

			\begin{itemize}
				\item PCIe access times $\mathcal{O}(\si{\micro\second})$.\footnotemark
				\item MATs recompiled $\mathcal{O}(\si{\second})$. Many rules $\implies$ batching.
				\item Thrift serde time $\mathcal{O}(\si{\milli\second})$.
			\end{itemize}

@@

			\alert{Meanwhile core-to-core on chip around \SI{100}{\nano\second} @ \SI{1.2}{\giga\hertz}.}
		\end{column}
		\begin{column}{0.5\linewidth}
			\begin{figure}
				\includegraphics[width=0.9\linewidth]{../plots/build/rte-timer/rte-times-logx}
				\caption{Netronome rule installation cost \\(1 table, \numrange{1}{65536} rules).}
			\end{figure}
		\end{column}
	\end{columns}

	\footcitetext{DBLP:conf/sigcomm/NeugebauerAZAL018}
\end{frame}

\begin{frame}{Quantisation}
	\begin{columns}
		\begin{column}{0.5\linewidth}
			\begin{figure}
				\includegraphics[width=0.9\linewidth]{../plots/build/marl-quant/accuracy-binary}
				\caption{DDoS throttle accuracy wrt. fixed-point Q$n$.}
			\end{figure}
		\end{column}
		\begin{column}{0.5\linewidth}
			No FPU... so use \alert{Q$n$ fixed-point}.
			\begin{itemize}
				\item For RL updates: need only shift, multiply, add.
				\item \alert{Policy exec only needs add}.
				\item \SI{16}{\bit} fraction has negligible loss.
			\end{itemize}
		\end{column}
	\end{columns}
\end{frame}

\begin{frame}{Dynamic Action Installation}
	\begin{itemize}
		\item Netronome w/ P4?
		\begin{itemize}
			\item MAT structure (DCFL) computed on host...
			\item So can't directly interop with P4.
			\item ...But we can have other actions/externs maintain their own lookup tables.
		\end{itemize}
		\item Tofino has features to make this more feasible.
		\begin{itemize}
			\item Action Profiles/Selects
		\end{itemize}
	\end{itemize}
\end{frame}

%\begin{frame}{What about higher port-density devices?}
%	content...
%\end{frame}

\begin{frame}{Execution costs (Q15.16)}
	\begin{columns}
		\begin{column}{0.4\linewidth}
			\begin{itemize}
				\item Single-threaded---still to be accelerated.
				\item Small? One tiling per memory tier.
				\item Max? 20-element state, 17 mixed-dim tilings, 8 per set.
				\item Versus use-case: \SI{330}{\micro\second} on commodity i7 (\SI{4.2}{\giga\hertz}).
			\end{itemize}
		\end{column}
		\begin{column}{0.6\linewidth}
			\centering
			\begin{tabular}{ l c c }
				\toprule
				Policy Size & Action (\si{\micro\second}) & Plus Update (\si{\micro\second}) \\ \midrule
				Small & \num{10.66} & \num{14.60} \\
				Max (DDoS) & \num{512.67} & \num{612.55} \\ \bottomrule
			\end{tabular}
		\end{column}
	\end{columns}
\end{frame}

\begin{frame}[standout]
	Takeaways:
	\begin{itemize}
		\item \alert{Online in-NIC RL is possible!}
		\item In-switch? Less so.
		\item Platform-specific, but similar design for SmartNIC hardware class.
		\item \alert{Work-in-progress}: end-to-end timing, training accuracy, other use-cases (AQM?), optimisation.
	\end{itemize}
	\alert{Questions?}
\end{frame}

>>>>

\usepackage{bm}
\usepackage{mathtools}

@@

\newcommand{\approach}{On Path Learning}
\newcommand{\approachshort}{OPaL}
\newcommand{\Coopfw}{\emph{CoOp}}
\newcommand{\coopfw}{\Coopfw}
\newcommand{\Indfw}{\emph{Ind}}
\newcommand{\indfw}{\Indfw}
\newcommand{\inring}{\textsc{In}}
\newcommand{\outring}{\textsc{Out}}

\makeatletter\let\expandableinput\@@input\makeatother
\newcommand{\cmark}{\ding{51}}%
\newcommand{\xmark}{\ding{55}}%

@@

\title{Revisiting the Classics: Online RL in the Programmable Dataplane}
\author{\vspace{-1em}\textbf{Kyle A. Simpson}, Dimitrios P. Pezaros\\
	\faEnvelopeO{} \href{mailto:k.simpson.1@research.gla.ac.uk}{\nolinkurl{k.simpson.1@research.gla.ac.uk}}\\
	\vspace{1em}\small{\faGithub{} \href{https://github.com/felixmcfelix}{FelixMcFelix} \hspace{0.5em} \faGlobe{} \url{https://mcfelix.me}}}
\institute{University of Glasgow}
\date{\nth{4} June, 2021}

\begin{document}
% title fun, including Org logos....
\begin{frame}
	\maketitle
	\begin{tikzpicture}[overlay, remember picture]
%		\node[above right=0.8cm and 0.9cm of current page.south west] (esnet-logo) {\includegraphics[width=2.75cm]{netlab-trim}};
%		\node[right=1cm of esnet-logo] {\adjincludegraphics[height=2cm,trim={0 {.4\height} 0 {.05\height}},clip]{uofg}};
		\node[above right=-0.1cm and 0.8cm of current page.south west] (uofg-logo) {\adjincludegraphics[height=2cm,trim={0 {.4\height} 0 {.05\height}},clip]{uofg}};
		\node[right=0.5cm of uofg-logo] {\includegraphics[width=2.75cm]{netlab-trim}};
	\end{tikzpicture}
\end{frame}

@@

@@

			\begin{itemize}
				\item PCIe access times $\mathcal{O}(\si{\micro\second})$\footnotemark
				\item Crossing VMs/vNFS has $\sim$\SI{10}{$\times$} higher cost\footnotemark
				\item MATs recompiled $\mathcal{O}(\si{\second})$. Many rules $\implies$ batching.
%				\item Thrift serde time $\mathcal{O}(\si{\milli\second})$.
			\end{itemize}

@@

			\alert{Meanwhile core-to-core on NFP around \SI{140}{\nano\second} @ \SI{1.2}{\giga\hertz}.}
		\end{column}
		\begin{column}{0.5\linewidth}
			\begin{figure}
				\includegraphics[width=0.9\linewidth]{../plots/build/rte-timer/rte-times-logx}
				\caption{Netronome rule installation cost \\(1 table, \numrange{1}{65536} rules).}
			\end{figure}
		\end{column}
	\end{columns}
	\setcounter{footnote}{3}
	\footcitetext{DBLP:conf/sigcomm/NeugebauerAZAL018}
	\setcounter{footnote}{4}
	\footcitetext{DBLP:journals/cm/CzivaP17}
\end{frame}

\begin{frame}{How do we bring online, in-NIC RL?}
	\begin{columns}
		\begin{column}{0.6\linewidth}
			\begin{figure}
				\includegraphics[width=\linewidth]{../paper/figures/arch-with-p4}
			\end{figure}
		\end{column}
		\begin{column}{0.4\linewidth}
			\begin{itemize}
				\item Classical RL built on tile-coding---\alert{online}.
				\item \alert{Fixed-point arithmetic}.
				\item Async wrt. datapath.
				\item Dynamic selection of last reward, trace info.
				\item \alert{Runtime configurable} (policy, size, application) from data/control-plane. \alert{Task independent.}
			\end{itemize}
		\end{column}
	\end{columns}
\end{frame}

\section{Background and Design}

\begin{frame}{Single-step (classical) RL---Sarsa}
	A simple explanation:\pause

	\begin{subequations}
		\begin{gather}
			\delta_t = \overbrace{R_{t+1} + \gamma \acval{S_{t+1}}{A_{t+1}}{\wvec{t}}}^{\mathclap{\substack{\text{New target value:}\\ \text{Reward} + \text{some of the next action's value}}}} - \underbrace{\acval{S_t}{A_t}{\wvec{t}}}_{\mathclap{\text{Current value estimate}}},\\\pause
			\bm{w}_{t+1} = \bm{w}_{t} + \overbrace{\alpha \delta_t}^{\mathclap{\text{Move a little bit of } \delta_t \text{ along...}}} \underbrace{\nabla{\acval{S_t}{A_t}{\wvec{t}}}}_{\mathclap{\text{...the policy's gradient}}},\pause
		\end{gather}%
	\end{subequations}

	Design implications?
\end{frame}

\begin{frame}{Tile Coding}
\centering
\resizebox{0.7\linewidth}{!}{
	\begin{tikzpicture}
		\node at (0.35,0.4){$s = \begin{pmatrix}
				0.7 \\
				0.3
			\end{pmatrix}$};
		\node at (0,-1.1) {$\bm{x}(s,\cdot) = \begin{Bmatrix}
				T_{1,9}, \\
				T_{2,5}, \\
				T_{\mathit{bias}}
			\end{Bmatrix}$};

		\node at (2.5,-0.5) {
			\begin{tikzpicture}
				\draw[step=0.5cm,color=uofgcobalt,opacity=0.7,shift={(0,0)},label=above:{Tiling 0}] (-0.5,-0.5) grid (1,1);
				\fill[uofgcobalt,opacity=0.5] (0.5,-0.5) rectangle (1,0);
				\node[color=uofgcobalt] at (0,1.2) {\footnotesize Tiling 1};

				\draw[step=0.5cm,color=uofgpumpkin,opacity=0.9,shift={(0.25,-0.25)},label=above:{Tiling 1}] (-0.5,-0.5) grid (1,1);
				\fill[uofgpumpkin,opacity=0.5,shift={(0.25,-0.25)}] (0,0) rectangle (0.5,0.5);
				\node[color=uofgpumpkin!50!uofgrust] at (0.25,-0.95) {\footnotesize Tiling 2};

				\node[circle, black, draw,
				fill, radius=0.5pt, inner sep=0pt,minimum size=1.5pt, label=above:{$s$}] at (0.625,-0.125) {};
				%			\filldraw (0.625,-0.125) circle[radius=1.5pt,label=above:{$s$}];

				\draw[->] (-0.25,-0.5)--(-0.25,0.85);
				\draw[->] (-0.25,-0.5)--(1.1,-0.5);

				\node at (1,-0.7) {\footnotesize 1};
				\node at (-0.4,0.75) {\footnotesize 1};
				\node at (-0.35,-0.6) {\footnotesize 0};
			\end{tikzpicture}
		};

	\end{tikzpicture}
}
\end{frame}

\begin{frame}{Tile Coding (ii)}
	\begin{columns}
		\begin{column}{0.6\linewidth}
			\centering
			\resizebox{0.8\linewidth}{!}{
				\begin{tikzpicture}
					\node at (0.35,0.4){$s = \begin{pmatrix}
							0.7 \\
							0.3
						\end{pmatrix}$};
					\node at (0,-1.1) {$\bm{x}(s,\cdot) = \begin{Bmatrix}
							T_{1,9}, \\
							T_{2,5}, \\
							T_{\mathit{bias}}
						\end{Bmatrix}$};

					\node at (2.5,-0.5) {
						\begin{tikzpicture}
							\draw[step=0.5cm,color=uofgcobalt,opacity=0.7,shift={(0,0)},label=above:{Tiling 0}] (-0.5,-0.5) grid (1,1);
							\fill[uofgcobalt,opacity=0.5] (0.5,-0.5) rectangle (1,0);
							\node[color=uofgcobalt] at (0,1.2) {\footnotesize Tiling 1};

							\draw[step=0.5cm,color=uofgpumpkin,opacity=0.9,shift={(0.25,-0.25)},label=above:{Tiling 1}] (-0.5,-0.5) grid (1,1);
							\fill[uofgpumpkin,opacity=0.5,shift={(0.25,-0.25)}] (0,0) rectangle (0.5,0.5);
							\node[color=uofgpumpkin!50!uofgrust] at (0.25,-0.95) {\footnotesize Tiling 2};

							\node[circle, black, draw,
							fill, radius=0.5pt, inner sep=0pt,minimum size=1.5pt, label=above:{$s$}] at (0.625,-0.125) {};
							%			\filldraw (0.625,-0.125) circle[radius=1.5pt,label=above:{$s$}];

							\draw[->] (-0.25,-0.5)--(-0.25,0.85);
							\draw[->] (-0.25,-0.5)--(1.1,-0.5);

							\node at (1,-0.7) {\footnotesize 1};
							\node at (-0.4,0.75) {\footnotesize 1};
							\node at (-0.35,-0.6) {\footnotesize 0};
						\end{tikzpicture}
					};

				\end{tikzpicture}
			}
		\end{column}
		\begin{column}{0.4\linewidth}
			\begin{itemize}[<+->]
				\item Operations? $+, -, \times, \div$
				\begin{itemize}[<+->]
					\item PoT tile width? $\div$ replaced w/ shift.
				\end{itemize}
				\item Tiling set: identical dimensions, different shifts
				\item Gradient for RL?
				\begin{itemize}[<+->]
					\item List of hit tiles.
				\end{itemize}
				\item Can be more complex...
			\end{itemize}
		\end{column}
	\end{columns}
\end{frame}
